---
title: "Penguins & NYCflights dataset manipultion and visuaization"
author: "Luke Geel"
date: "10/19/2020"
output:
  pdf_document: default
  html_document: default
---
```{r}
library(tidyverse)
library(palmerpenguins)
```
Question 1
A. What are the percentages of missing values in bill_length_mm, bill_depth_mm, sex? (recall the percent() function in scales package)
```{r}
d <- nrow(penguins)
a <- nrow(penguins%>%filter(is.na(bill_length_mm)))
b <- nrow(penguins%>%filter(is.na(bill_depth_mm)))
c <- nrow(penguins%>%filter(is.na(sex)))
"% of missing bill length:" 
a/d *100
"% of missing bill depth:"
b/d*100
"% of missing bill length:"
c/d*100
```
B. What are the means and medians of the body mass for male and female penguins of each species? (Remove the NAs in the sex first)
```{r}
penguins_no_NA <- penguins %>% filter(!is.na(sex))
penguins_no_NA %>% group_by(species,sex) %>%summarize(mean_body_mass = mean(body_mass_g), median_body_mass = median(body_mass_g))
```
C. Make the following plot. (x label is removed, and the unit of y axis is changed to kilogram (kg))
```{r}
penguins_in_kg <- penguins %>%filter(!is.na(sex))%>%mutate(body_mass_kg = (body_mass_g/1000))
ggplot(data = penguins_in_kg)+
  geom_boxplot(mapping = aes(x=sex, y=body_mass_kg))+
  facet_wrap(. ~species, nrow =1)+
  labs(title = "Average body mass", y="body mass (in kg)")
```
D. Make the following plot.
```{r}
penguins%>%filter(!is.na(sex))%>%group_by(sex, year,species)%>%mutate(body_mass_kg = (body_mass_g/1000))%>%mutate(mean = mean(body_mass_kg))%>%
ggplot(aes(x=year, y=mean))+
  geom_line(mapping = aes(linetype = sex))+
  geom_point(aes(shape = sex))+
  facet_wrap(~species)+
  labs(title = "Average body mass over time", y="body mass (in kg)", x="year")+
  scale_y_continuous(breaks=seq(0,5.5,0.5))+
  scale_x_continuous(breaks=seq(2007.0,2009.0,1), labels=c("07","08","09"))

```
E. Make the following two plots. (The units of the variables are changed)
```{r}
penguins_in_kg <- penguins %>%filter(!is.na(sex))%>%mutate(body_mass_kg = (body_mass_g/1000), flipper_length_cm = flipper_length_mm/10, bill_length_cm = bill_length_mm/10)

ggplot(data = penguins_in_kg)+
  geom_smooth(mapping = aes(linetype = sex, x=flipper_length_cm,y=body_mass_kg))+
  facet_wrap(~species, nrow=1)+
  theme()+
  labs(title = "Flipper length vs. Body mass", y="body mass (in KG)", x="flipper length (in cm)")

ggplot(data = penguins_in_kg)+
  geom_smooth(mapping = aes(linetype = sex, x=flipper_length_cm,y=bill_length_cm))+
  facet_wrap(~species, nrow=1)+
  theme()+
  labs(title = "Flipper length vs. Bill length", y="bill length (in cm)", x="flipper length (in cm)")
```
Question 2
A. Find out the most frequent destination airports for all domestic flights departing from New York City (in terms of total number of flights).
```{r}
library(nycflights13)
flights %>% filter(!is.na(flight)) %>% group_by(dest) %>% summarize(n_flights = n_distinct(flight)) %>% arrange(desc(n_flights))


```
B. Passengers are often frustrated when their flight departs late, but are not as annoyed if, in the end, pilots can make up some time during the flight. Extract the flights data in October and calculate the “gained time in the air” as the difference between arr_delay and dep_delay. Make a scatterplot of dep_delay vs. the “gained time” and add a smooth line.
```{r}
flights %>% filter(month == 10)%>%mutate(gained_time_in_air = arr_delay - dep_delay)%>%
  ggplot(aes(y=dep_delay, x=gained_time_in_air))+geom_point()+geom_smooth()+labs(title = "Depature delay vs Gained time", y="depature delay", x="gained time")
```
C. The “gained time” defined above may not be reasonable because long distance flight may be more likely to gain more. Let’s define the “gained time per hour” and draw a histogram of this variable. What are the mean and median of this variable?
```{r}
flights%>%filter(month == 10)%>%filter(!is.na(arr_delay) & !is.na(dep_delay))%>%mutate(gained_time_per_hour = (arr_delay - dep_delay)/air_time)%>%
  ggplot(aes(x=gained_time_per_hour))+geom_histogram(bins = 75)

flights %>% filter(month == 10 & !is.na(arr_delay) & !is.na(dep_delay)) %>% mutate(gained_time_per_hour = (arr_delay - dep_delay)/air_time)%>%summarize(mean=mean(gained_time_per_hour), median = median(gained_time_per_hour))


```
D. Plot the following bar chart on the number of flights took off every 3 hours. You need to
1. remove cancelled flights
2. create a variable which equals 0 if the flight took off between 0 and 3 am, equals 1 if took off between 3 and 6 am…
3. make the bar chart
4. change the labels
```{r}
flights%>%filter(!is.na(dep_time),!is.na(arr_time))
attach(flights)
flights$hourcat[hour >= 0 & hour <3 ] <- "0-3"
flights$hourcat[hour >= 3 & hour <6] <- "3-6"
flights$hourcat[hour >= 6 & hour <9] <- "6-9"
flights$hourcat[hour >= 9 & hour <12] <- "9-12"
flights$hourcat[hour >= 12 & hour <15] <- "12-15"
flights$hourcat[hour >= 15 & hour <18] <- "15-18"
flights$hourcat[hour >= 18 & hour <21] <- "18-21"
flights$hourcat[hour >= 21] <- "21-24"
flights%>%mutate(hourcat = fct_relevel(hourcat, "0-3", "3-6", "6-9", "9-12", "12-15", "15-18", "18-21", "21-24"))%>%
ggplot()+
  geom_bar(aes(x= hourcat))+
  labs(title = "", y="", x="Hour")

```
Question 3
A. Select the babies born in 2010 and used by no less than 10000 babies.
(optional) Obtain the names that are used by no less than 10000 boys and 10000 girls.
```{r}
library(babynames)
babynames%>%filter(year == 2010, n >= 10000)

babynames%>% group_by(sex)%>%
  filter(n>10000)%>%
  arrange(sex)


```
B. Find out the longest names (in terms of the number of letters) in 2010. (str_length() calculates the number of letters of a string).
```{r}
babynames%>%filter(year == 2010)%>%mutate(letters_in_name = str_length(name))%>%arrange(desc(letters_in_name))
```
C. Make a time series plot (line plot) of the number of babies having a long name (>=10 letters). Below is an example.

```{r}
babynames%>%mutate(letters_in_name = str_length(name))%>%group_by(year)%>%filter(letters_in_name >= 10)%>%mutate(total_count = sum(n))%>%
ggplot()+geom_line(aes(x=year,y=total_count))+
  labs(x="year",y="count")
```
D. Use proportions (number of long name babies/ total number of babies in that year) instead of counts for the previous question.
```{r}
babynames%>%
  summarize(n=sum(n))

babynames%>%
  mutate(num_lett_name = str_length(name))%>%
  filter(num_lett_name>=10)%>%
  group_by(year)%>%
  summarize(num_of_prop = (sum(n)/348120517)*100)%>%
  ggplot(aes(x=year,y=num_of_prop*100))+
  geom_line()+
  labs(x="year",y="precent")
``` 

Question 4
```{r}
toy <- tibble(
  x = c(4, 2, 2, 3, 3, 1, 3, 4, 5),
  group = c("c", "b", "a", "c", "b", "a", "a", "b", "c")
)
```
A. Suppose we would like to sort x in an increasing order within each group like this
Does the following code work?
toy %>%
  group_by(group) %>%
  arrange(x)
No, this code does not work as it sorts by the group first, rather than sort by x.
This code now works:
toy %>%
  group_by(group) %>%
  arrange(group,x)
```{r}
toy %>%
  group_by(group) %>%
  arrange(group,x)
```
B. Suppose we would like to subtract the within group mean from each x. For example, the first observation in toy is in group c, with x equal to 4. Then we want to have 4 minus the mean of the three values in group c, (4+3+5)/3. Does the following code work? Make it work using group_by().
toy %>%
  mutate(x_new = x - mean(x))
```{r}
toy %>%
  group_by(group)%>%
  mutate(x_new = x - mean(x))%>% 
  arrange(group,x)

```
  
